---
title: "Data Idea"
output: html_document
date: "2022-08-16"
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r}
library(tidyverse)
library(leaflet)
library(sf)
```

## Data Cleaning

```{r}
mortality <- read_csv(here::here("data/share-of-deaths-by-cause.csv"))
glimpse(mortality)

world_map <- read_sf(here::here("data/world-administrative-boundaries/world-administrative-boundaries.shp")) %>%
  st_zm(drop = TRUE)

require(gapminder)
new <- gapminder %>%
  mutate(country = tolower(as.character(country)))

countries <- unique(new$country)

mortality_sf <- mortality %>%
  filter(tolower(Entity) %in% countries) %>%
  group_by(Code) %>%
  left_join(world_map, by = c("Code" = "iso3")) %>%
  ungroup()

NCD <- c("Cardiovascular diseases", "Neoplasms", "Chronic respiratory diseases", "Diabetes mellitus")
```

```{r}
clean_mortality <- mortality_sf %>%
  select(Entity,
         Code,
         Year,
         `Deaths - Cardiovascular diseases - Sex: Both - Age: All Ages (Percent)`,
         `Deaths - Neoplasms - Sex: Both - Age: All Ages (Percent)`,
         `Deaths - Maternal disorders - Sex: Both - Age: All Ages (Percent)`,
         `Deaths - Chronic respiratory diseases - Sex: Both - Age: All Ages (Percent)`,
         `Deaths - Digestive diseases - Sex: Both - Age: All Ages (Percent)`,
         `Deaths - Diabetes mellitus - Sex: Both - Age: All Ages (Percent)`,
         `Deaths - Lower respiratory infections - Sex: Both - Age: All Ages (Percent)`,
         `Deaths - Neonatal disorders - Sex: Both - Age: All Ages (Percent)`,
         `Deaths - Diarrheal diseases - Sex: Both - Age: All Ages (Percent)`,
         `Deaths - Cirrhosis and other chronic liver diseases - Sex: Both - Age: All Ages (Percent)`,
         `Deaths - Tuberculosis - Sex: Both - Age: All Ages (Percent)`,
         `Deaths - Chronic kidney disease - Sex: Both - Age: All Ages (Percent)`,
         `Deaths - Alzheimer's disease and other dementias - Sex: Both - Age: All Ages (Percent)`,
         `Deaths - Parkinson's disease - Sex: Both - Age: All Ages (Percent)`,
         `Deaths - HIV/AIDS - Sex: Both - Age: All Ages (Percent)`,
         `Deaths - Acute hepatitis - Sex: Both - Age: All Ages (Percent)`,
         `Deaths - Malaria - Sex: Both - Age: All Ages (Percent)`,
         `Deaths - Nutritional deficiencies - Sex: Both - Age: All Ages (Percent)`,
         `Deaths - Meningitis - Sex: Both - Age: All Ages (Percent)`,
         `Deaths - Protein-energy malnutrition - Sex: Both - Age: All Ages (Percent)`,
         `Deaths - Enteric infections - Sex: Both - Age: All Ages (Percent)`,
         continent,
         geometry) %>%
  pivot_longer(-c(Entity, Code, Year, continent, geometry),
               values_to = "Mortality_rate_(%)") %>%
  mutate(Disease = str_replace_all(name, c("Deaths - " = "", " - Sex: Both - Age: All Ages \\(Percent\\)" = "")),
         NCD = ifelse(Disease %in% NCD, "Yes", "No")) %>%
  select(-name) %>%
  relocate(`Mortality_rate_(%)`, .after = NCD) %>%
  relocate(geometry, .after = `Mortality_rate_(%)`)
```


## Motivation, Purpose

- Target Audience: Healthcare Professional
- Purposes: The purpose of this app is to spread awareness on the diseases around the world
- Motivation: With the COVID-19 pandemic and new diseases, may shadow the other existing diseases, hence this app is to provide information and emphasis on the ongoing diseases that have caused death for more than a decade.


## Analysis

```{r}
clean_mortality %>%
  filter(continent == "Africa",
         Year == 2019) %>%
  group_by(Disease) %>%
  summarise(Mortality_rate = mean(`Mortality_rate_(%)`), .groups = "drop") %>%
  arrange(-Mortality_rate)

clean_mortality %>%
  filter(continent == "Americas",
         Year == 2019) %>%
  group_by(Disease) %>%
  summarise(Mortality_rate = mean(`Mortality_rate_(%)`), .groups = "drop") %>%
  arrange(-Mortality_rate)

clean_mortality %>%
  filter(continent == "Europe",
         Year == 2019) %>%
  group_by(Disease) %>%
  summarise(Mortality_rate = mean(`Mortality_rate_(%)`), .groups = "drop") %>%
  arrange(-Mortality_rate)

clean_mortality %>%
  filter(continent == "Oceania",
         Year == 2019) %>%
  group_by(Disease) %>%
  summarise(Mortality_rate = mean(`Mortality_rate_(%)`), .groups = "drop") %>%
  arrange(-Mortality_rate)

clean_mortality %>%
  filter(continent == "Asia",
         Year == 2019) %>%
  group_by(Disease) %>%
  summarise(Mortality_rate = mean(`Mortality_rate_(%)`), .groups = "drop") %>%
  arrange(-Mortality_rate)
```   

```{r}
clean_mortality %>%
  filter(Entity == "Australia",
         Year == 2019) %>%
  arrange(-`Mortality_rate_(%)`)
```

```{r}

```

