---
title: "Data Idea"
output: html_document
date: "2022-08-16"
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r}
library(tidyverse)
library(leaflet)
library(sf)
library(plotly)
library(knitr)
library(kableExtra)
```

## Data Cleaning

```{r}
mortality <- read_csv(here::here("data/share-of-deaths-by-cause.csv"))
glimpse(mortality)

world_map <- read_sf(here::here("data/world-administrative-boundaries/world-administrative-boundaries.shp")) %>%
  st_zm(drop = TRUE)

new <- world_map %>%
  mutate(iso3 = tolower(as.character(iso3)))

countries <- unique(new$iso3)

mortality_sf <- mortality %>%
  filter(tolower(Code) %in% countries) %>%
  group_by(Code) %>%
  right_join(world_map, by = c("Code" = "iso3")) %>%
  ungroup()
```

```{r}
clean_mortality <- mortality_sf %>%
  select(Entity,
         Code,
         Year,
         `Deaths - Cardiovascular diseases - Sex: Both - Age: All Ages (Percent)`,
         `Deaths - Neoplasms - Sex: Both - Age: All Ages (Percent)`,
         `Deaths - Maternal disorders - Sex: Both - Age: All Ages (Percent)`,
         `Deaths - Chronic respiratory diseases - Sex: Both - Age: All Ages (Percent)`,
         `Deaths - Digestive diseases - Sex: Both - Age: All Ages (Percent)`,
         `Deaths - Diabetes mellitus - Sex: Both - Age: All Ages (Percent)`,
         `Deaths - Lower respiratory infections - Sex: Both - Age: All Ages (Percent)`,
         `Deaths - Neonatal disorders - Sex: Both - Age: All Ages (Percent)`,
         `Deaths - Diarrheal diseases - Sex: Both - Age: All Ages (Percent)`,
         `Deaths - Cirrhosis and other chronic liver diseases - Sex: Both - Age: All Ages (Percent)`,
         `Deaths - Tuberculosis - Sex: Both - Age: All Ages (Percent)`,
         `Deaths - Chronic kidney disease - Sex: Both - Age: All Ages (Percent)`,
         `Deaths - Alzheimer's disease and other dementias - Sex: Both - Age: All Ages (Percent)`,
         `Deaths - Parkinson's disease - Sex: Both - Age: All Ages (Percent)`,
         `Deaths - HIV/AIDS - Sex: Both - Age: All Ages (Percent)`,
         `Deaths - Acute hepatitis - Sex: Both - Age: All Ages (Percent)`,
         `Deaths - Malaria - Sex: Both - Age: All Ages (Percent)`,
         `Deaths - Nutritional deficiencies - Sex: Both - Age: All Ages (Percent)`,
         `Deaths - Meningitis - Sex: Both - Age: All Ages (Percent)`,
         `Deaths - Protein-energy malnutrition - Sex: Both - Age: All Ages (Percent)`,
         `Deaths - Enteric infections - Sex: Both - Age: All Ages (Percent)`,
         continent,
         geometry) %>%
  na.omit() %>%
  pivot_longer(-c(Entity, Code, Year, continent, geometry),
               values_to = "Mortality_rate_(%)") %>%
  mutate(Disease = str_replace_all(name, c("Deaths - " = "", " - Sex: Both - Age: All Ages \\(Percent\\)" = "")),
         `Mortality_rate_(%)` = round(`Mortality_rate_(%)`, 4)) %>%
  select(-name) %>%
  relocate(geometry, .after = `Mortality_rate_(%)`) 
```


## Motivation, Purpose

- Target Audience: Healthcare Professional
- Purposes: The purpose of this app is to spread awareness on the diseases around the world
- Motivation: With the COVID-19 pandemic and new diseases, may shadow the other existing diseases, hence this app is to provide information and emphasis on the ongoing diseases that have caused death for more than a decade.


## Analysis

```{r}
clean_mortality %>%
  filter(continent == "Africa",
         Year == 2019) %>%
  group_by(Disease) %>%
  summarise(Mortality_rate = mean(`Mortality_rate_(%)`), .groups = "drop") %>%
  arrange(-Mortality_rate)

clean_mortality %>%
  filter(continent == "Americas",
         Year == 2019) %>%
  group_by(Disease) %>%
  summarise(Mortality_rate = mean(`Mortality_rate_(%)`), .groups = "drop") %>%
  arrange(-Mortality_rate)

clean_mortality %>%
  filter(continent == "Europe",
         Year == 2019) %>%
  group_by(Disease) %>%
  summarise(Mortality_rate = mean(`Mortality_rate_(%)`), .groups = "drop") %>%
  arrange(-Mortality_rate)

clean_mortality %>%
  filter(continent == "Oceania",
         Year == 2019) %>%
  group_by(Disease) %>%
  summarise(Mortality_rate = mean(`Mortality_rate_(%)`), .groups = "drop") %>%
  arrange(-Mortality_rate)

clean_mortality %>%
  filter(continent == "Asia",
         Year == 2019) %>%
  group_by(Disease) %>%
  summarise(Mortality_rate = mean(`Mortality_rate_(%)`), .groups = "drop") %>%
  arrange(-Mortality_rate) %>%
  select(-Mortality_rate)


clean_mortality %>%
  filter(Entity == "World Bank High Income",
         Year == 2019) %>%
  group_by(Disease) %>%
  summarise(Mortality_rate = mean(`Mortality_rate_(%)`), .groups = "drop") %>%
  arrange(-Mortality_rate)

clean_mortality %>%
  filter(Entity == "World Bank Low Income",
         Year == 2019) %>%
  group_by(Disease) %>%
  summarise(Mortality_rate = mean(`Mortality_rate_(%)`), .groups = "drop") %>%
  arrange(-Mortality_rate)
```   

```{r}
clean_mortality_highest <- clean_mortality %>%
  group_by(Entity, Year) %>%
  mutate(Highest = max(`Mortality_rate_(%)`),
         Cause = ifelse(Highest == `Mortality_rate_(%)`, "Yes", "No")) %>%
  filter(Cause == "Yes") %>%
  select(-c(Cause, Highest))

clean_mortality_highest <- st_as_sf(clean_mortality_highest) %>%
  filter(Year == 2019)

disease_ordered <- clean_mortality_highest %>%
  group_by(Disease) %>%
  summarise(n = n()) %>%
  arrange(-n) %>%
  pull(Disease)

clean_mortality_highest$Disease <- ordered(clean_mortality_highest$Disease, levels = disease_ordered)
```

```{r}
library(tmap)
library(RColorBrewer)
tmap_mode("view")
tm_shape(clean_mortality_highest %>%
           filter(continent == "Americas")) +
  tm_polygons(col = "#ABBAD7") +
  tm_basemap(leaflet::providers$CartoDB.VoyagerNoLabels) +
  tm_view(set.view = c(-78,35,1.5))


ggplot(clean_mortality_highest %>%
         filter(continent == "Europe")) +
  geom_sf()

```

```{r}
tm_shape(clean_mortality_highest) +
  tm_polygons(col = "Disease", palette = "-YlOrRd") +
  tm_basemap(leaflet::providers$CartoDB.VoyagerNoLabels)
```



```{r}
p <- clean_mortality %>%
  filter(Entity == "Australia",
         Disease == "Cardiovascular diseases") %>%
  ggplot(aes(x = Year, y = `Mortality_rate_(%)`)) +
  geom_line(color = "red") +
  geom_point(color = "red") +
  theme_bw() +
  labs(x = "Year", y = "Mortality Rate (Percentage)",
       title = "Mortality Rate Over Time") +
  theme(axis.text.x = element_text(angle = 50, vjust = 1, hjust = 1),
        plot.title = element_text(face = "bold",
                                  margin = margin(10, 0, 10, 0),
                                  size = 14))

ggplotly(clean_mortality %>%
  filter(Entity == "Australia",
         Disease == "Cardiovascular diseases") %>%
  ggplot(aes(x = Year, y = `Mortality_rate_(%)`)) +
  geom_line(color = "red") +
  geom_point(color = "red") +
  theme_bw() +
  labs(x = "Year", y = "Mortality Rate (Percentage)",
       title = "Mortality Rate Over Time") +
  theme(axis.text.x = element_text(angle = 50, vjust = 1, hjust = 1),
        plot.title = element_text(face = "bold",
                                  margin = margin(10, 0, 10, 0),
                                  size = 14)))

clean_mortality %>%
  filter(continent == "Asia",
         Year == 2019) %>%
  group_by(Disease) %>%
  summarise(Mortality_rate = mean(`Mortality_rate_(%)`), .groups = "drop") %>%
  arrange(-Mortality_rate) %>%
  select(-Mortality_rate) %>%
  mutate(Rank = paste(1:21,".")) %>%
  relocate(Disease, .after = Rank) %>%
  head(10) %>%
  kable() %>%
  kable_styling(bootstrap_options = c("striped", "hover"), full_width = FALSE)
```


```{r}
ordered <- clean_mortality %>%
  filter(continent == "Asia",
         Year == 2019) %>%
  group_by(Disease) %>%
  summarise(Mortality_rate = mean(`Mortality_rate_(%)`), .groups = "drop") %>%
  arrange(Mortality_rate) %>%
  pull(Disease)


bar <- clean_mortality %>%
  filter(continent == "Asia",
         Year == 2019)

bar$Disease <- ordered(bar$Disease, levels = ordered)

ggplotly(bar %>%
  group_by(Disease) %>%
  summarise(Mortality_rate = mean(`Mortality_rate_(%)`), .groups = "drop") %>%
  arrange(-Mortality_rate) %>%
  ggplot(aes(x = Disease, y = `Mortality_rate`)) +
  geom_bar(stat = "identity", fill = "red") +
  theme_bw() +
  labs(x = "Disease Name", y = "Mortality Rate (Percentage)") +
  coord_flip())
```


